---
output: github_document
---

# Distribution Generalization under Dense Distributional Shifts

## How to install

1. The [devtools](https://github.com/hadley/devtools) package has to be installed. You can install it using  `install.packages("devtools")`.
2. The latest development version can then be installied using `devtools::install_github("yujinj/dlm")`.

```{r}
#devtools::install_github("yujinj/dlm", force=TRUE)
library(dlm)
```

## Usage: Applications on GTEx Data

In this example, we use the [GTEx V6](https://www.gtexportal.org/home/downloads/adult-gtex/qtl) gene expression data. The data has already been preprocessed and normalized. We simply filter out genes to only include those observed in all tissues and then randomly sample 5000 genes to reduce the size of the data set. The preprocessed data set is provided in the package as 'GTEx_data'.

```{r, include = FALSE}
load("~/Documents/dlm/data/GTEx_data.rda")
```

First, one has to choose $L$ different test functions. Then, define a single test function that takes a data set and applies your choice of $L$ test functions. If the sample size of the data set is $n$, the output should be $n \times L$. In this GTEx example, we choose $L = 1000$ test functions as the products of gene-expressions for randomly selected $L$ gene-pairs.

```{r}
set.seed(1000)
L = 1000
names = colnames(GTEx_data[[1]])
names1 <- sample(names, size=L, replace=FALSE)
names2 <- sample(names, size=L, replace=FALSE)
test.function <- function(data){return(data[,names1] * data[,names2])}
phi <- function(data){return(data[,names1] * data[,names2])}
```

We obtain the correlation matrix of 1000 test functions and check that they are mostly weakly correlated. For example, 90% of the correlations are within the range: [-0.059,  0.059].

```{r}
phi_matrix = lapply(GTEx_data, FUN=phi)
phi_matrix = do.call(rbind, phi_matrix)
cor_phi = cor(phi_matrix)
quantile(cor_phi[row(cor_phi)!=col(cor_phi)], c(0.05, 0.95))
```

### Example in the Paper

The target tissue is Brain_Cortex. Other available tissues are Adipose_Subcutaneous, Adipose_Visceral_Omentum, Brain_Frontal_Cortex_BA9, and Brain_Cerebellum.

```{r}
dlm.fit = dlm(Brain_Cortex ~ Adipose_Subcutaneous + 
                             Adipose_Visceral_Omentum +
                             Brain_Frontal_Cortex_BA9 +
                             Brain_Cerebellum,
              test.function = phi, 
              data = GTEx_data, 
              whitening = FALSE)
```

```{r}
dlm.fit
```

We present a pair plot where each row and column represents one tissue, and each dot represents the mean of the test function. In our case, the mean of the test function corresponds to the covariance of a randomly selected gene-pair. We can observe some linear relationships between tissues, which are reflected in our 'dlm' results.

```{R}
pairs(dlm.fit$X, cex = 0.2, labels = c("Brain\nCortex", "Adipose\nSubcutaneous", "Adipose\nVisceral\nOmentum", "Brain\nFrontal\nCortex", "Brain\nCerebellum"))
```

To evaluate the appropriateness of the fitted distribution shift model, we plot a distributional residual plot and distributional QQ plot. 

```{r, fig.width=5, fig.height=8}
par(mfrow = c(2, 1))
plot(dlm.fit$lm.fit, which = c(1, 2))
```      

### More Examples

We randomly select 6 tissues: Artery_Aorta, Vagina, Esophagus_Gastroesophageal_Junction, Esophagus_Muscularis, Muscle_Skeletal, Skin_Not_Sun_Exposed_Suprapubic. 

```{r}
set.seed(1000)
sampled_tissues = sample(names(GTEx_data), 6)
sampled_tissues
```

Let the target tissue be Artery_Aorta.

```{r}
dlm.fit = dlm(Artery_Aorta ~ Vagina + 
                             Esophagus_Gastroesophageal_Junction + 
                             Esophagus_Muscularis + 
                             Muscle_Skeletal +
                             Skin_Not_Sun_Exposed_Suprapubic,
              test.function = phi, 
              data = GTEx_data, 
              whitening = FALSE)
```

```{r}
dlm.fit
```

```{R}
pairs(dlm.fit$X, cex = 0.2, labels = gsub("_", "\n", sampled_tissues))
```

Let the target tissue be Esophagus_Muscularis.

```{r}
dlm.fit = dlm(Esophagus_Muscularis ~ Artery_Aorta +
                                     Vagina + 
                                     Esophagus_Gastroesophageal_Junction + 
                                     Muscle_Skeletal +
                                     Skin_Not_Sun_Exposed_Suprapubic,
              test.function = phi, 
              data = GTEx_data, 
              whitening = FALSE)
```

```{r}
dlm.fit
```
